---
title: Tidymodels
author: Björn Koneswarakantha
date: '2020-11-12'
slug: tidymodels
categories:
  - R
  - modelling
tags:
  - R
  - tidymodels
  - modelling
summary: A preview on the tidymodels meta package
thumbnailImagePosition : left
thumbnailImage: https://avatars0.githubusercontent.com/u/29100987?s=400&v=4
editor_options: 
  chunk_output_type: console
output:
  blogdown::html_page:
    toc: true
---

<br></br>
<br></br>

```{r short_logo, echo = FALSE}
blogdown::shortcode("image"
                    , classes = 'center'
                    , src = 'https://avatars0.githubusercontent.com/u/29100987?s=400&v=4'
                    , thumbnail = 'https://avatars0.githubusercontent.com/u/29100987?s=400&v=4'
                    , `thumbnail-width` = '360px'
                    , `thumbnail-height` = '360px'
                    , target = 'https://github.com/tidymodels/tidymodels'
                    )

```

<br></br>
<br></br>

```{r include = FALSE}
knitr::opts_chunk$set(warning = F)
```

# Introduction

Two years ago I took the tidymodels packages for a test drive. Now some of the packages at experimental stage have been released on CRAN and are ready to be added to my exemplatory modelling flow. The biggest change is that we will use `parsnip`, `dials`, `tune` and `workflows` instead of `caret`.

We will still try to organize all steps in a modeling dataframe and use `purrr` and `furrr` maping functions whenever possible.

# Modeling

We will fit the following regression models to the Boston Housing Data Set

- xgbTree
- lm
- randomForest
- MARS
- CART tree

For tuning we will use a randomized parameter search in a 5-fold cross validation

We will use the following packages:
-`recipes`
-`resample`
-`dials`
-`tune`
-`parsnip`
-`yardstick`
-`workflows`

```{r}
suppressPackageStartupMessages( library('mlbench') )
suppressPackageStartupMessages( library('tidyverse') )
suppressPackageStartupMessages( library('recipes') )
suppressPackageStartupMessages( library('caret') )
suppressPackageStartupMessages( library('Hmisc') )
suppressPackageStartupMessages( library('xgboost') )
suppressPackageStartupMessages( library('parsnip') )
suppressPackageStartupMessages( library('dials') )
suppressPackageStartupMessages( library('tune') )
suppressPackageStartupMessages( library('workflows') )
suppressPackageStartupMessages( library('xgboost') )
suppressPackageStartupMessages( library('doFuture') )
suppressPackageStartupMessages( library('parallel') )
suppressPackageStartupMessages( library('easyalluvial') )


# ggplot default theme
theme_set(theme_minimal())

```



## Data

```{r}
data('BostonHousing')
df = as_tibble( BostonHousing )
summary(df)
```


## Preprocessing with recipe


**Note we are intentionally standardizing the response variable since the unit of lstat is irrelevant for this demo**

We will 

- Yeo Johnson Transform
- Scale
- Center
- remove co-correlating variables (threshold 0.5)
- dummy encode

```{r}
rec = recipe(df, lstat ~ . )

summary(rec)
```

```{r}
rec = rec %>%
  step_scale(all_numeric(), - all_outcomes()) %>%
  step_center(all_numeric(), - all_outcomes()) %>%
  step_YeoJohnson(all_numeric()) %>%
  step_corr(all_numeric(), - all_outcomes(), threshold = 0.5 ) %>%
  step_dummy(all_nominal())
  

```

### Summary Recipe

```{r}
prep_rec = prep(rec, df)
prep_rec
df_prep = bake(prep_rec, df )
```

## Resampling with rsample


```{r}
rs = rsample::vfold_cv(df, v = 5)

rsample::pretty.vfold_cv(rs)
```


## Modelling with parsnip

- xgbTree
- lm
- lm with regularisation
- randomForest
- MARS
- CART tree

### Modelling Dataframe with Workflows

```{r}

get_wflow_all_params_set_to_tune <- function(.f, engine, mode, recipe) {
  
  arg_names <- formals(.f) %>%
    names()
  
  # remove arguments if they do not have a dials function
  arg_names <- intersect(arg_names, ls("package:dials"))
  
  arg_values <- rep(list(tune()), length(arg_names))
  
  arg_comb <- setNames(arg_values, arg_names)
  
  arg_comb$mode <- mode
  
  m <- do.call(.f, arg_comb) %>%
    parsnip::set_engine(engine)
  
  wf <- workflows::workflow() %>%
    workflows::add_model(m) %>%
    workflows::add_recipe(recipe)
  
  return(wf)
}

finalize_unkowns <- function(wf, rs) {
  data <- rs$splits[[1]]$data
  pset <- parameters(wf)
  predictors <- workflows::.fit_pre(wf, data)$pre$mold$predictors
  pset$object <- purrr::map(pset$object, dials::finalize, x = predictors)
  return(pset)
}

df_m <- tibble(
    .f_parsnip = c(parsnip::decision_tree, parsnip::boost_tree, parsnip::rand_forest, parsnip::mars, parsnip::linear_reg),
    mode = "regression",
    engine = c("rpart", "xgboost", "randomForest", "earth", "glmnet") 
  ) %>%
  mutate(wf = pmap(list(.f_parsnip, engine, mode), get_wflow_all_params_set_to_tune, rec),
         pset = map(wf, finalize_unkowns, rs))


df_m

```

#### Tuning

##### Parallel Processing

```{r}
all_cores <- parallel::detectCores(logical = FALSE)

library(doFuture)
registerDoFuture()
cl <- parallel::makeCluster(all_cores)
plan(cluster, workers = cl)
```


##### Bayes 

```{r cache = TRUE}

df_m_tune_bayes <- df_m %>%
  mutate(
     tune = map2(
       wf,
       pset,
       function(x,y) tune_bayes(
         object = x,
         rs,
         iter = 10,
         param_info = y,
         metrics = yardstick::metric_set(
           yardstick::rmse,
           yardstick::rsq,
           yardstick::mape,
           yardstick::smape,
           yardstick::rpd,
           yardstick::mase
         ),
         control = control_bayes(save_pred = TRUE)
      )
    )
  )

df_m_tune_bayes

```

#### Random

```{r cache = TRUE}
df_m_tune_rnd <- df_m %>%
  # filter(! engine %in% c("xgboost")) %>%
  # filter(engine == "earth") %>%
  mutate(
     tune = map2(
       wf,
       pset,
       function(x,y) tune_grid(
         x,
         rs,
         iter = 10,
         param_info = y,
         metrics = yardstick::metric_set(
           yardstick::rmse,
           yardstick::rsq,
           yardstick::mape,
           yardstick::smape,
           yardstick::mase
        ),
        control = control_grid(save_pred = TRUE)
      )
    )
  )

df_m_tune_rnd
```


#### Finalize

```{r}

possibly_best_param <- purrr::possibly(tune::select_best, otherwise = NA)
possibly_best_metric <- purrr::possibly(tune::show_best, otherwise = NA)
possibly_final_wf <- purrr::possibly(tune::finalize_workflow, otherwise = NA)
possibly_fit <- purrr::possibly(parsnip::fit, otherwise = NA)
possibly_pulled_m <- purrr::possibly(workflows::pull_workflow_fit, otherwise = NA)
possibly_imp <- purrr::possibly(vip::vi_model, otherwise = NA)

df_m_tune <- bind_rows(
  mutate(df_m_tune_bayes, tune_meth = "bayes"),
  mutate(df_m_tune_rnd, tune_meth = "rnd")
  )

df_m_final <- df_m_tune %>%
  mutate( best_metrics = map(tune, possibly_best_metric, metric = "rmse"),
          best_param = map(tune, possibly_best_param, metric = "rmse"),
          final_wf = map2(wf, best_param, possibly_final_wf),
          final_fit = map(final_wf, possibly_fit, df),
          m = map(final_fit, possibly_pulled_m),
          imp = map(m, possibly_imp))

df_m_final

```

### Plot Model Metrics

```{r fig.height=7}

get_fold_metrics <- function(tu, best_param) {
  tu %>%
    unnest(.metrics) %>%
    inner_join(best_param) %>%
    select(id, .metric, .estimate)
}

df_plot <- df_m_final %>%
  filter(! is.na(best_param)) %>%
  mutate(fold_metrics = map2(tune, best_param, get_fold_metrics)) %>%
  unnest(fold_metrics) %>%
  mutate(method = paste0(engine, "_", tune_meth)) %>%
  select(method, id, .metric, .estimate) %>%
  group_by(method, .metric) %>%
  mutate(me = mean(.estimate),
         se = sd(.estimate) / sqrt(n()))

df_plot

df_1se = df_plot %>%
  select(method, .metric, me, se) %>%
  distinct() %>%
  mutate(  ymin = me - se
          , ymax = me + se ) %>%
  group_by(.metric) %>%
  mutate( rnk = rank(me, ties.method = 'first')
          , rnk_desc = rank( desc(me), ties.method = 'first')
          ) %>%
  rename( best_method = method ) %>%
  filter( (rnk == 1 & .metric != 'rsq') | (.metric == 'rsq' & rnk_desc == 1) )

df_1se

len = df_plot$method %>%
  n_distinct()

col_class = RColorBrewer::brewer.pal('Greys', n = 9) %>% rev()
col_folds = RColorBrewer::brewer.pal('Dark2', n = 8) %>%
  easyalluvial::palette_filter(greens = F, greys = F)

pal = c( col_class[2], col_folds[1:5], col_class[4], col_class[6] )

df_plot %>%
  left_join( select(df_1se, .metric, ymin, ymax, best_method) ) %>%
  mutate( classification = case_when( best_method == method ~ 'best'
                                      , me >= ymin & me <= ymax ~ 'in'
                                      , T ~ 'out' )
          ) %>%
  ggplot( aes(method, .estimate) ) +
    geom_rect( aes(ymin = ymin, ymax = ymax )
               , xmin = 0, xmax = len
               , fill = col_class[7]
               , alpha = 0.05
               ) +
    geom_line( aes( group = id, color = id)
               , size = .5, alpha = 1 ) +
    stat_summary( aes( color = classification)
                  , size = 1
                  , fun.data = function(x) mean_se( x, mult = 1) ) +
    scale_color_manual( values = pal ) +
    theme( legend.position = 'none', axis.text.x = element_text(angle = 90, vjust = 0.5)) +
    labs( y = '', x = '', caption = 'grey area: 1SE range of best model'
        , title = 'CV Performance Metrics + SE') +
    facet_wrap(~.metric, scales = 'free_y', ncol = 1) 
```

### Plot Predictions

```{r}
get_test_predictions <- function(tu, best_param, data = df_prep) {
  tu %>%
    unnest(.predictions) %>%
    inner_join(best_param) %>%
    select(id, .pred, .row) %>%
    mutate(.obs = df_prep$lstat[.row])
}

df_plot <- df_m_final %>%
  filter(! is.na(best_param)) %>%
  mutate(test_preds = map2(tune, best_param, get_test_predictions)) %>%
  unnest(test_preds) %>%
  mutate(method = paste0(engine, "_", tune_meth)) 

df_plot %>%
  ggplot(aes(.pred, .obs)) +
    geom_point(alpha = 0.2) +
    geom_smooth() +
    facet_grid(tune_meth ~ engine) +
    geom_abline(slope = 1, color = "tomato") +
    theme(aspect.ratio = 1)

```

### Plot Partial Dependence Alluvial Plots

```{r}
df_pdap <- df_m_final %>%
  mutate(is_lgl = map_lgl(final_fit, ~ is.logical(.))) %>%
  filter(! is_lgl) %>%
  mutate(pdap = map(final_fit, easyalluvial::alluvial_model_response_parsnip, df, degree = 4, bins = 5, resp_var = "lstat", method = "pdp", parallel = TRUE))

wf <- df_m_final$final_fit[[1]]
wf

p <- easyalluvial::alluvial_model_response_parsnip(wf, df, method = "pdp", parallel = TRUE, resp_var = "lstat")

predict(wf, new_data = df)

wf <- df_m_final$final_fit[[1]]

predict(wf, new_data = df)

m <- decision_tree() %>%
  set_engine("rpart")

wf <- workflow() %>%
  add_model(m) %>%
  add_recipe(prep_rec) %>%
  fit(df)

predict(wf, new_data = df)


p <- easyalluvial::alluvial_model_response_parsnip(m, df_prep, method = "pdp", parallel = TRUE, )

grid <- p %>%
  easyalluvial::add_marginal_histograms(data_input = df_prep) %>%
  easyalluvial::add_imp_plot(p = p, data_input = df_prep, )
```

