---
title: Quality Control with R - Notes
author: Björn Koneswarakantha
date: '2021-08-27'
slug: qcwithr
categories:
  - Rtraining
tags:
  - R
  - Rtraining
summary: Quality Control with R - Notes
thumbnailImagePosition : left
thumbnailImage: https://media.springernature.com/w306/springer-static/cover-hires/book/978-3-319-24046-6
editor_options: 
  chunk_output_type: console
output:
  blogdown::html_page:
    toc: true
---

<br></br>
<br></br>

```{r, echo = FALSE}
blogdown::shortcode("image"
                    , classes = 'center'
                    , src = 'https://media.springernature.com/w306/springer-static/cover-hires/book/978-3-319-24046-6'
                    , thumbnail = 'https://media.springernature.com/w306/springer-static/cover-hires/book/978-3-319-24046-6'
                    , `thumbnail-width` = '400px'
                    , `thumbnail-height` = '600px'
                    , target = 'https://www.qualitycontrolwithr.com/'
                    )

```


> Quality Control with R. An ISO Standards Approach (Cano EL, Moguerza JM and Prieto M, 2015).

<br></br>
<br></br>


```{r include = FALSE}
knitr::opts_chunk$set(warning = F)
```


# R packages

`qcc` seems to be the most popular and best maintained QC package for control charts.

```{r load}
suppressPackageStartupMessages(library(tidyverse))
# we use qcc 3.0 which is not released on CRAN yet
# devtools::install_github("luca-scr/qcc", build = TRUE, build_opts = c("--no-resave-data", "--no-manual"))
library(qcc)
library(AcceptanceSampling)
library(SixSigma)
library(DiagrammeR)
```


# ISO standard QC Approach

ISO standards are defined by the International Organisation for Standardisation by technical committees using a process that is similar to peer-review. There are a lot of ISO standards covering Quality Control.

# QC tools

There are 7 general quality Ishikava tools, each of them is covered by numerous ISO documents.

- Cause and Effect Diagram
- Check sheet
- Control chart
- Histogram
- Pareto Chart
- Scatter Diagram
- Stratification
  * Flow charts

## Cause and Effect Diagrams

The effect is usually undesired. As causes for a given effect 
for processes of the producing industry the ISO standard is 5Ms+E

- Manpower
- Materials
- Machines
- Methods
- Measurements
- Environment

Creatively I would adapt this for data-driven business processes to:

- Manpower
- Events
- IT System
- Methods (Statistics, DevOps, Code base, Manual Processes)
- Measurements of Events (Data Entry, Data Coding)
- Environment (Culture, External Events, Regulations)

Specifically for the reporting of Adverse Drug Reactions in the form of individual case safety reports (ICSR) to Health Authorities (HA).

```{r, fig.width=10, fig.height = 7}
qcc::causeEffectDiagram(
  cause = list(
    Manpower = c(
      "Medical Reviewer",
      "Global Process Owner",
      "Local Process Owner",
      "IT Developer",
      "Database Manager",
      "Auditer"
    ),
    Event = c(
      "ICSR Quality",
      "ICSR Detectability",
      "ICSR Quantity",
      "ICSR Complexity"
    ),
    `IT Systems` = c(
      "Availability",
      "Complexity",
      "Performance",
      "Quality"
    ),
    Methods = c(
      "Medial Review Process\n(Data Guides)",
      "Decisioning\n(Submission Rules)",
      "Risk Management",
      "Quality Mangement System",
      "Change Management"
    ),
    Measurements = c(
      "Data Entry",
      "Medical Coding"
    ),
    Environment = c(
      "HA Regulations",
      "Local Cultures",
      "Company Culture",
      "Epidemiology",
      "Local Health Systems"
    )
  ),
  effect = "ICSR not\nreported\nto HA"
)
```

## Check Sheets

Serve to generate process related data, should be part of a general data collection plan. Will be automated for most processes.

- Audit Trail Data
- Quality Evidence
- IT Logs

Classically serves to attribute each effect (ICSR not reported to HA) to any of the suspected causes in the cause and effect diagram.

## Control Charts (Shewhart Charts)

Process Monitoring Charts, can be used to track whether process KPI are in or out of control.

```{r}
data("pistonrings")
df <- as_tibble(pistonrings)

df_g1 <- df %>%
  group_by(sample) %>%
  mutate(rwn = row_number()) %>%
  filter(rwn == 1)

qcc_shew <- qcc(
  df$diameter,
  type = "xbar.one",
  rules = c(1, 2, 3, 4)
)

qcc_shew
str(qcc_shew)
plot(qcc_shew)

qcc_shew$violations
qcc_shew$rules
```

`qcc` uses the [Western Electric Rules](https://en.wikipedia.org/wiki/Western_Electric_rules) to flag out of control values.

- One point plots outside 3-sigma control limits.
- Two of three consecutive points plot beyond a 2-sigma limit.
- Four of five consecutive points plot beyond a 1-sigma limit.
- Eight consecutive points plot on one side of the center line.

These rules do not appear to be customizable, it seems to me that [Nelson Rules](https://en.wikipedia.org/wiki/Nelson_rules) are
a bit more complete and cover more non-random scenarios such as trends and oscillation.

The control chart type depends on the data type and its distribution. There is a decision tree for the chart selection.

```{r}
DiagrammeR::mermaid(
  "graph TB
    A(Select Control Chart)-->B{data type}
    B --> |Discrete| C{Distribution}
    B --> |Continuous| D{Sample Size}
    C --> |Poisson| E{Sample Size}
    C --> |Binominal| F{Frequency}
    E --> |constant| G[c-chart]
    E --> |variable| H[u-chart]
    F --> |count| I[np-chart]
    F --> |proportion| J[p-chart]
    D --> |n = 1| K[x.one + R chart]
    D --> |n = 2-10| L[x + R chart]
    D --> |n > 10| M[x + S chart]
"
)
```

## Histogram

self explanatory, explore distribution of continuous variable.

```{r}
ggplot(df_g1) +
  geom_histogram(aes(diameter))
```

## Pareto Chart

Plot Frequency of effect causes, apply [80:20 rule](https://en.wikipedia.org/wiki/Pareto_principle)

```{r}
defect <- c(80, 27, 66, 94, 33)
names(defect) <- c("HA Regulations", "Risk Management", "Data Entry", "Change Management", "IT System Performance")
qcc_pareto <- qcc::paretoChart(defect)
str(qcc_pareto)

plot(qcc_pareto) +
  theme(axis.text.x = element_text(angle = 90))
```

## Scatter Plots

- validate established cause and effect relationships
- multivariate control diagrams

## Stratification

- identify relevant process subgroups by which to stratify the analysis
- use of process flow charts

# Control Charts

- ideally have even sample sizes to obtain even upper and lower control limits, although `qcc` allow you to create control charts with uneven sample sizes but oc plots cannot be created, thus it is difficult to make good estimates about appropriate sample sizes.
- plot in pairs to monitor different summary statistics like x-bar chart (means) and range chart (R chart) as a measure for variance
- use Shewart's formula to estimate control limits
- phase I determine a calibration period during which the process is in control and determine control limits
- phase II monitor process for out of range values

## Continuous Grouped Variables

We are going to measure a continuous variable over time such as the thickness of a board.

Controls Charts use the **population** standard deviation to flag out of control data points.

Determine **population** standard deviation:

- determine a calibration period during which the process is in control (set n)
- at each time point measure samples and take the mean
- according to the [central limit theorem](https://en.wikipedia.org/wiki/Central_limit_theorem) the distribution of the means will be normally distributed
- measure population standard deviation of the means `sqrt((n-1)/n) * sd(x)`
- in statistical process control this is considered to be biased
- instead the range and the shewart constant (d2/c4) which depends on the number of measurements (n) are used by the Shewart's formula to estimate control limits

Determine sample size:
- ARL is the average run length (number of consecutive samples) needed to get a point that is outside the upper and lower control limits. For a normal distribution the ARL is 370. So every 370 measurements we will get a false positive signal.
- The power of a control chart depends on the sample size. The operator characteristics (OC) plots the probabilty of a false negative (type two error) against the process shift in standard deviations for several sample sizes (n). For a given shift in process deviation we can also determine the probability *beta* for a false negative using the OC curve.
- ARL = 1 / (1 - *beta*)

### X-bar Chart

- Monitors the mean
- uses d2 Shewart's constants

```{r}
df

# pivot from long to wide
qcc_gr <- qcc::qccGroups(df, diameter, sample)

qcc_gr

q1 <- qcc(
  qcc_gr[1:25, ],
  type = "xbar",
  newdata = qcc_gr[26:nrow(qcc_gr), ],
  rules = c(1, 2, 3, 4)
)

q1

plot(q1)

q1_oc <- qcc::ocCurves(q1)

q1_oc

plot(q1_oc)

arl_n5 <- q1_oc$ARL %>%
  as_tibble(rownames = "shift_stdev") %>%
  filter(shift_stdev == "1.0") %>%
  pull(`5`)

arl_n5
```

So for a change in average diameter by `r q1$std.dev` we need in average `r arl_n5` runs to detect a change in the mean statistic using 5 samples.

### R Chart

- monitors process stability by means of the sample ranges
- uses d3 and d2 Shewart's constants
- appropriate for sample sizes < 8

```{r}
q2 <- qcc(
  qcc_gr[1:25, ],
  type = "R",
  newdata = qcc_gr[26:nrow(qcc_gr), ],
  rules = c(1, 2, 3, 4)
)

q2

plot(q2)

q2_oc <- qcc::ocCurves(q2)

# throws error, https://github.com/luca-scr/qcc/issues/31
# plot(q2_oc)

q2_oc

arl_n5 <- q2_oc$ARL %>%
  as_tibble(rownames = "scale_multiplier") %>%
  filter(scale_multiplier == "1.5") %>%
  pull(`5`)

arl_n5
```

So for a change in range by 50% need in average `r arl_n5` runs using 5 samples.


### S Chart

- monitors process stability by means of standard deviations
- uses c4 Shewart's constant
- appropriate for sample sizes >= 8

```{r}
q3 <- qcc(
  qcc_gr[1:25, ],
  type = "S",
  newdata = qcc_gr[26:nrow(qcc_gr), ],
  rules = c(1, 2, 3, 4)
)

q3

plot(q3)

q3_oc <- qcc::ocCurves(q3)

# throws error, https://github.com/luca-scr/qcc/issues/31
# plot(q3_oc)

q3_oc

arl_n5 <- q3_oc$ARL %>%
  as_tibble(rownames = "scale_multiplier") %>%
  filter(scale_multiplier == "1.5") %>%
  pull(`5`)
```

So for a change in range by 50% need in average `r arl_n5` runs using 5 samples.


## Continuous Non-Grouped Variables

If a process is described best by a continuous one at a time measurement we cannot use sampling

- central limit theorem does not apply
- normality needs to be confirmed
- if possible use power transformation to normalize
- accompany x-bar chart with moving range chart

### X-bar chart

- uses the moving range to calculate global standard deviation
- takes Shewart's constant d2 for n = 2

```{r}
# airplane paint viscosity, one batch takes hours to produce and no rational subgroups can be formed
data("viscosity")
df_vis <- viscosity

df_vis

q4 <- qcc(
  df_vis$viscosity[df_vis$trial],
  type = "xbar.one",
  newdata = df_vis$viscosity[! df_vis$trial],
  rules = c(1, 2, 3, 4)
)

q4

plot(q4)
```

### R Chart

- uses the range between a given and the next measurement

```{r}

vis_mat <- df_vis %>%
  mutate(lead = lead(viscosity)) %>%
  select(viscosity, lead) %>%
  filter(! is.na(lead)) %>%
  as.matrix()


q5 <- qcc(
  vis_mat[df_vis$trial[1:nrow(df_vis)-1], ],
  type = "R",
  newdata = vis_mat[! df_vis$trial[1:nrow(df_vis)-1], ]
)

q5

plot(q5)
```

## Discrete Measurements

### p and np Charts

- measure frequency or relative frequency
- p is for relative frequency 
- np for frequency

```{r}
data("orangejuice")
df_oj <- orangejuice

q6 <- qcc(
  df_oj$D[df_oj$trial],
  df_oj$size[df_oj$trial],
  type = "p",
  newdata = df_oj$D[! df_oj$trial],
  newsizes = df_oj$size[! df_oj$trial],
  rules = c(1, 2, 3, 4)
)

q6

plot(q6)

q6_oc <- qcc::ocCurves(q6)

q6_oc

plot(q6_oc)

plot(q6_oc, what = "ARL")

q6a <- qcc(
  df_oj$D[df_oj$trial],
  df_oj$size[df_oj$trial],
  type = "np",
  newdata = df_oj$D[! df_oj$trial],
  newsizes = df_oj$size[! df_oj$trial],
  rules = c(1, 2, 3, 4)
)

q6a

plot(q6a)

```

### c chart

- for count of events that cannot be attributed to a sample size
- examples: nonconformities per day, defects per m^2 of fabric or flaws per metal plate
- unit should stay constant

```{r}
library(SixSigma)

ss.data.thickness2 %>%
  head(15)

# remove uninspected items
df_flaws <- ss.data.thickness2 %>%
  as_tibble() %>%
  filter(! is.na(flaws))

df_flaws

q7 <- qcc(
  df_flaws$flaws,
  type = "c",
  rules = c(1, 2, 3, 4)
)

q7

plot(q7)

q7_oc <- qcc::ocCurves(q7)

q7_oc

plot(q7_oc, what = "ARL") +
  coord_cartesian(ylim = c(0, 10))
```


### u chart

- average number of defects from a varying number of samples
- for example: 18 flaws from 3 metal plates, then 3 flaws in 5 metal plates

```{r}
df_gr_flaws <- df_flaws %>%
  group_by(ushift) %>%
  summarise(n_plates_per_shift = n(),
            n_flaws_per_shift = sum(flaws))

df_gr_flaws

qu8 <- qcc(
  df_gr_flaws$n_flaws_per_shift,
  type = "u",
  sizes = df_gr_flaws$n_plates_per_shift,
  rules = c(1, 2, 3, 4)
)

qu8

plot(qu8)

# OC curves need constant sample sizes
# qu8_oc <- qcc::ocCurves(qu8)
```

## Capability/Performance Indices

- measure whether consumer specified upper and lower specification limits (LSL/USL) are compatible with process control limits (LCL/UCL)
- performance measures long-term variability and capacity short-term variability
- C = (USL - LSL) / 6sigma
- index below 1 is unsufficient, 1.33-167 is outstanding and 2 refers to six sigma quality
- adjusted index variations account for processes that are not perfectly centered

Cp:   capability index
Cp_l: lower index
Cp_u: upper index
Cp_k: minimum of lower and upper index
Cpm : Tagauchi index (controls for not centered processes)

```{r}

q1

q1_pc <- qcc::processCapability(q1, spec.limits = c(73.975, 74.025))

q1_pc

plot(q1_pc)

q1_pc <- qcc::processCapability(q1, spec.limits = c(73.95, 74.05))

q1_pc

plot(q1_pc)

```

## CUSUM Chart

- cumulative sum chart
- two lines one below zero and one above zero and a center which is the expected value. 
- positive deviation of the process from the central value results in an uptick of both lines
- negative deviation of the process from the central value results in a down tick of both lines
- both lines never cross the central line
- supposedly more sensitive than regular control charts

```{r}
qcc_gr

q1 <- qcc(
  qcc_gr[1:25, ],
  type = "xbar",
  newdata = qcc_gr[26:nrow(qcc_gr), ],
  rules = c(1, 2, 3, 4)
)

plot(q1)

q9 <- qcc::cusum(
  qcc_gr[1:25, ],
  newdata = qcc_gr[26:nrow(qcc_gr), ]
)

plot(q9)
```

## EWMA Chart

- exponentially weighted moving average chart
- zj = gamma * xj + (1 - gamma) * zj-1
- if the smoothing parameter gamma is 0.2 a given value for x is 20% present value and 80% past values
- `+` indicates the actual values
- supposed to work well for not normal values

```{r}
q10 <- qcc::ewma(
  qcc_gr[1:25, ],
  newdata = qcc_gr[26:nrow(qcc_gr), ]
)

plot(q10)
```

# Sampling/Stratification

Naturally samples need to be representative of the population. If there are known subgroups it is important to adapt a stratified sampling strategy.

[free ebook](https://bookdown.org/lawson/an_introduction_to_acceptance_sampling_and_spc_with_r26/)

## Attribute Acceptance Sampling

- attribute: defective / not defective
- producer wants a high probability of acceptance for a lot with a low defective fraction
- producers and consumer agree on acceptable quality level (AQL)
- alpha (producers risk) is the probability with which a lot that passes AQL is rejected
- consumer wants a high probability of rejection for a lot with a high defective ratio
- producers and consumer agree on lot tolerance percent defective (LTPD)
- beta (consumers risk) is the probability with which a lot that does not pass LTPD is accepted

Requirements:

- AQL: 0.06
- alpha: 0.05
- LTPD: 0.16
- beta: 0.10

```{r}
library(AcceptanceSampling)

plan <- AcceptanceSampling::find.plan(
  PRP = c(0.06, 0.95),
  CRP = c(0.16, 0.10)
)

plan

oc <- AcceptanceSampling::OC2c(plan$n, plan$c, plan$r)

plot(oc, xlim = c(0, 0.25))
```

take a sample of `r plan$n` and reject when `r plan$r` items are defective

## Variable Acceptance Sampling

specify:

- upper specification limit (USL)

and or:

- lower specification limit (LSL)

```{r}
plan2 <- AcceptanceSampling::find.plan(
  PRP = c(0.06, 0.95),
  CRP = c(0.16, 0.10),
  type = "normal",
  s.type = "known"
)

plan2

```

proceed:

- take a sample of `r plan2$n`  
- compute the mean  
- compute LSL/USL - mean(sample) / sd  
- compare accept if value above is greater `r plan2$k`  

```{r}
plan3 <- AcceptanceSampling::find.plan(
  PRP = c(0.06, 0.95),
  CRP = c(0.16, 0.10),
  type = "normal",
  s.type = "unknown"
)

plan3
```

proceed:

- take a sample of `r plan3$n`  
- compute the mean  
- compute LSL/USL - mean(sample) / sd(sample)  
- compare accept if value above is greater `r plan3$k`

Note that sample size has increased because we do not know population sd and need to estimate it from the sample.
